---
title: "DGM-mouseFMRI"
author: "Simon Schwab"
date: "12 June 2018"
output: html_notebook
---

## Install required packages
```{r}
# install.packages("rmarkdown")
# install.packages("DGM")
# install.packages("cowplot")
# install.packages("ppcor")
# install.packages("abind")
# install.packages("testit")
```

## Libraries
```{r message=FALSE}
library(ggplot2)
library(DGM)
library(reshape2)
library(cowplot)
library(ppcor)
library(abind)
library(testit)
library(data.table)
```

## Variables
```{r}
INFO = 'mouseFMRI'
Nn=18
N=34
Nt=900
TR=1
PATH_HOME = '/home/simon'
PATH_PROJ = file.path(PATH_HOME, 'Data', 'DGM-mouseFMRI') # Project path
PATH_RESULTS = file.path(PATH_PROJ, 'results')
PATH_FIGS = file.path(PATH_PROJ, 'figures')

PATH_NET = file.path(PATH_HOME, 'Drive', 'DGM_mouseFMRI')
PATH_TS = file.path(PATH_HOME, 'Drive', 'mouseFMRI')

br = Nn - 0.5 # break for figures
```

## Load mouse info
```{r}
tmp = read.table(file = file.path(PATH_RESULTS, 'info.txt'), colClasses = "character")
mydata = data.frame(file=tmp$V1, id=tmp$V2, virus=tmp$V3, active=tmp$V4)
print(mydata)

rois = read.table(file = file.path(PATH_RESULTS, 'labels.txt'), colClasses = "character")
mylabels = rois$V3

is.ctrl = mydata$virus=='Control'
is.active = mydata$active=='2'
N.contrl = 7
N.dreadd = 10

#         baseline active
# DREAD     n=10   n=10
# Ctrl      n=7    n=7
```

## Load time series data
```{r}
f = list.files(PATH_TS, pattern = '*.txt')
# read data
ts = array(NA, dim=c(Nt, Nn, N))
for (s in 1:N) {
  d = as.matrix(read.table(file.path(PATH_TS, f[s]), header = F))
  ts[,,s] = scaleTs(d[1:Nt,1:Nn]) # in case some subjects have more volumes
}
```

## Plot example time series
```{r, fig.height=3, fig.width=10}
set.seed(1980)
s = sample(N, 4)
n = sample(Nn, 3)
idx=1:100

p = list()
c = 1
for (i in s) {
  d = ts[,n,i]
  colnames(d)=mylabels[n]
  d_ = melt(d[idx,])
  d_$Var2=as.factor(d_$Var2)
  p[[c]] = ggplot(d_, aes(x = Var1, y = value, group=Var2, color=Var2)) + 
    geom_line() + theme_minimal() + ggtitle(f[i])
  c = c + 1
}

plot_grid(plotlist = p, ncol = 2, nrow = 2, rel_widths = c(1, 1))
```

## Run DGM on HCP cluster

## Load DGM networks
```{r eval=FALSE, include=FALSE}

subj=list()
for (s in 1:N) {
  print(sprintf("Loading mouse dataset %s", mydata$file[s]))
  subj[[s]] = read.subject(path=PATH_NET, id=sprintf("%s_%s", substr(mydata$file[s],18,22), INFO), Nn, modelStore = F)
}

dgm.all = dgm.group(subj)
dgm.contrl.base = dgm.group(subj[is.ctrl &  !is.active])
dgm.contrl.actv = dgm.group(subj[is.ctrl &  is.active])
dgm.dreadd.base = dgm.group(subj[!is.ctrl & !is.active])
dgm.dreadd.actv = dgm.group(subj[!is.ctrl & is.active])

save(dgm.all, dgm.contrl.base, dgm.contrl.actv, dgm.dreadd.base, dgm.dreadd.actv, file = file.path(PATH_RESULTS, sprintf('DGM-%s.RData', INFO)), compress = T)
```
### Load from RData
```{r}
load(file.path(PATH_RESULTS, sprintf('DGM-%s.RData', INFO)))
```

## Discount factors
```{r fig.height=4, fig.width=7}
set.seed(1980) # because jitter will always create a slightly different image

lim = c(0.95, 1)
p1 = ggplot(melt(dgm.contrl.base$df_), aes(x=as.factor(Var2), y=value)) +
  geom_point(shape=1, color="gray50", size=1, position = position_jitter(width = NULL, height = 0)) +
  geom_boxplot(width=0.6) + theme(axis.text = element_text(size=8)) +
  ggtitle(sprintf("controls baseline", N)) + ylab("df") + xlab("node") + ylim(lim)

p2 = ggplot(melt(dgm.contrl.actv$df_), aes(x=as.factor(Var2), y=value)) +
  geom_point(shape=1, color="gray50", size=1, position = position_jitter(width = NULL, height = 0)) +
  geom_boxplot(width=0.6) + theme(axis.text = element_text(size=8)) +
  ggtitle(sprintf("controls active", N)) + ylab("df") + xlab("node") + ylim(lim)

p3 = ggplot(melt(dgm.dreadd.base$df_), aes(x=as.factor(Var2), y=value)) +
  geom_point(shape=1, color="gray50", size=1, position = position_jitter(width = NULL, height = 0)) +
  geom_boxplot(width=0.6) + theme(axis.text = element_text(size=8)) +
  ggtitle(sprintf("DREADD baseline", N)) + ylab("df") + xlab("node") + ylim(lim)

p4 = ggplot(melt(dgm.dreadd.actv$df_), aes(x=as.factor(Var2), y=value)) +
  geom_point(shape=1, color="gray50", size=1, position = position_jitter(width = NULL, height = 0)) +
  geom_boxplot(width=0.6) + theme(axis.text = element_text(size=8)) +
  ggtitle(sprintf("DREADD active", N)) + ylab("df") + xlab("node") + ylim(lim)

plot_grid(p1, p2, p3, p4, ncol = 2, nrow = 2)
```
### DF Summary stats
```{r, fig.height=5.2, fig.width=6.5}
x=rbind(summary(c(dgm.contrl.base$df_)),
        summary(c(dgm.contrl.actv$df_)),
        summary(c(dgm.dreadd.base$df_)),
        summary(c(dgm.dreadd.actv$df_)))
rownames(x)=c("controls baseline", "controls active", "DREADD baseline", "DREADD active")
print(x)
```

### Proportion of time-varying fits vs. stationary networks (df = 1)
```{r}
x = array(NA, dim = c(1,4))
x[1,] = c(sum(dgm.contrl.base$df_ < 1) / (sum(is.ctrl & !is.active)*Nn),
          sum(dgm.contrl.actv$df_ < 1) / (sum(is.ctrl & is.active)*Nn),
          sum(dgm.dreadd.base$df_ < 1) / (sum(!is.ctrl & is.active)*Nn),
          sum(dgm.dreadd.actv$df_ < 1) / (sum(!is.ctrl & !is.active)*Nn))
colnames(x) = c("C base", "C active", "D base", "D active")
print(x)
```

## Stats
```{r}
stats.contrl.base = binom.nettest(dgm.contrl.base$am, alter = "greater", fdr = 0.05)
stats.contrl.actv = binom.nettest(dgm.contrl.actv$am, alter = "greater", fdr = 0.05)

stats.dreadd.base = binom.nettest(dgm.dreadd.base$am, alter = "greater", fdr = 0.05)
stats.dreadd.actv = binom.nettest(dgm.dreadd.actv$am, alter = "greater", fdr = 0.05)

stats.prop.contrl = prop.nettest(x1 = apply(dgm.contrl.actv$am, c(1,2), sum),
                                 n1 = sum(is.ctrl & is.active),
                                 x2 = apply(dgm.contrl.base$am, c(1,2), sum),
                                 n2 = sum(is.ctrl & !is.active),
                                 fdr = 0.05, alpha = 0.05)

stats.prop.dreadd = prop.nettest(x1 = apply(dgm.dreadd.actv$am, c(1,2), sum),
                                 n1 = sum(!is.ctrl & is.active),
                                 x2 = apply(dgm.dreadd.base$am, c(1,2), sum),
                                 n2 = sum(!is.ctrl & !is.active),
                                 fdr = 0.05, alpha = 0.05)

stats.prop.base = prop.nettest(x1 = apply(dgm.dreadd.base$am, c(1,2), sum),
                               n1 = sum(!is.ctrl & !is.active),
                               x2 = apply(dgm.contrl.base$am, c(1,2), sum),
                               n2 = sum(is.ctrl & !is.active),
                               fdr = 0.05, alpha = 0.05)

stats.prop.actv = prop.nettest(x1 = apply(dgm.dreadd.actv$am, c(1,2), sum),
                               n1 = sum(!is.ctrl & is.active),
                               x2 = apply(dgm.contrl.actv$am, c(1,2), sum),
                               n2 = sum(is.ctrl & is.active),
                               fdr = 0.05, alpha = 0.05)
```

## Calculate delta active - control
```{r}
delta.ctrl = array(NA, dim=c(Nn, Nn, N.contrl))
for (i in 1:N.contrl) {
  delta.ctrl[,,i] = dgm.contrl.actv$am[,,i] - dgm.contrl.base$am[,,i]
}

# stats.delta.ctrl = binom.nettest(delta.ctrl, fdr = .05)
```

## Plot network structures
### Determine limits
```{r}
x = c(stats.contrl.actv$adj - stats.contrl.base$adj,
      stats.dreadd.actv$adj - stats.dreadd.base$adj, 
      stats.contrl.base$adj - stats.dreadd.base$adj,
      stats.contrl.actv$adj - stats.dreadd.actv$adj)
mylim=max(abs(x))
mylim=c(-mylim, mylim)
```

##  Modeling a binomial outcome with one or more explanatory variables
For example motor, thalamus [2,17]
```{r}
logReg.P = array(NA, dim=c(Nn, Nn))
logReg.zval = array(NA, dim=c(Nn, Nn))

for (i in 1:Nn) {
  for (j in 1:Nn) {
    a = as.logical(dgm.all$am[i,j,])
    tmp = summary(glm(formula = a ~ is.active * !is.ctrl, family = binomial))
    logReg.P[i,j] = tmp$coefficients[4,4]
    logReg.zval[i,j] = tmp$coefficients[4,3]
  }
}

# p.adjust(logReg.P[!diag(Nn)], method = "fdr")
```

## Plot Network
```{r fig.height=6, fig.width=10, message=FALSE}
p1 = gplotMat(stats.contrl.base$adj, title = "controls (baseline)", barWidth = 0.3,
              nodeLabels = mylabels, xAngle = 90, axisTextSize = 7, colMapLabel = "%") +
  scale_x_continuous(breaks = 0.5:br, labels = mylabels)

p2 = gplotMat(stats.contrl.actv$adj, title = "controls (active)", barWidth = 0.3,
              nodeLabels = mylabels, xAngle = 90, axisTextSize = 7, colMapLabel = "%") +
  scale_x_continuous(breaks = 0.5:br, labels = mylabels)

p3 = gplotMat(logReg.zval*(logReg.P < 0.05), title = "DREADD*active (p < .05, uc.)", barWidth = 0.3, lim = c(-3,3), gradient = c("blue", "white", "red"),
              nodeLabels = mylabels, xAngle = 90, axisTextSize = 7, colMapLabel = "z") +
  scale_x_continuous(breaks = 0.5:br, labels = mylabels)

p4 = gplotMat(stats.dreadd.base$adj, title = "DREADD (baseline)", barWidth = 0.3,
              nodeLabels = mylabels, xAngle = 90, axisTextSize = 7, colMapLabel = "%") +
  scale_x_continuous(breaks = 0.5:br, labels = mylabels)

p5 = gplotMat(stats.dreadd.actv$adj, title = "DREADD (active)", barWidth = 0.3,
              nodeLabels = mylabels, xAngle = 90, axisTextSize = 7, colMapLabel = "%") +
  scale_x_continuous(breaks = 0.5:br, labels = mylabels)

plot_grid(p1, p2, p3, p4, p5, ncol = 3, nrow = 2, rel_widths = c(1, 1))
ggsave(path = PATH_FIGS, sprintf('DGM_%s-Proportions.png', INFO))
```



